An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-28 Revision/response.tex

\noindent
\textbf{Response:}

\section*{Response to Review 1}

%%%%%%%%%%%%%%%%%%%%%% W1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{weakness}
		\textit{Although the authors justify the synthetic attribute generation, real-world applications may involve more complex attribute semantics and correlations, which might affect generalizability.}
	\end{weakness}
}

\noindent
\textbf{Response:} I really appreciate your suggestion. Current hybrid search algorithms don't explicitly model or leverage complex semantic relationships, such as those between vectors and their attributes, or among different attributes. However, our evaluation covered a range of realistic scenarios, including multi-attribute setting, diverse distributions and selectivity, to assess the algorithms' generalization capabilities (though we used synthetic attributes).

To demonstrate that the choice between synthetic and real attributes has little impact on algorithm performance, we conducted an additional set of experiments. We used the YouTube-Audio dataset for vectors and compared results with both synthetic and real attributes (video classification). As shown in Figure~\ref{fig:compare}, the experimental outcomes validated our hypothesis.

@@

%%%%%%%%%%%%%%%%%%%%%% D2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{Section 4.2.1: It is important to run the experiments on larger datasets, e.g. 100 M vectors, or 1B bigann. Otherwise, differences will not show.
			Also, build time parameters such as Radius and Length which impact the Recall need to be considered.
			Also, approaches for partitioning and merging for building very large indexes need to be considered.
			Please also document whether quantization is included in index build time.
			There is no explanation why Vamana is not included in Figure 7.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for your insightful comments.

\textbf{Regarding dataset scale and scalability:}

We conducted experiments on a 100M-scale deep dataset and analyzed the performance of each algorithm on the large dataset.\textbf{ Please refer to MR.R7.}

@@

\textcolor{blue}{"The fused table is Table1: Comparison of AF-ANN and RF-ANN search algorithms."}

%%%%%%%%%%%%%%%%%%%%%% D4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{Section 4.3.
			It is not explained, what the Range attributes are, or how many attributes. I assume 1 attribute.
			The index construction times are vastly different in Fig 12 a. This calls for a much more detailed explanation.
			Also, Milvus seems to be missing in Fig 12 c. Why?
		}
	\end{detailed}
}

\noindent
\textbf{Response:} Thank you very much for your careful review and valuable suggestions. We greatly appreciate your feedback and sincerely apologize for any lack of clarity in our paper.

Regarding the issue of range attributes, we indeed used only one range attribute for evaluation. At present, most existing range filtering algorithms only support filtering on a single attribute. This is a major limitation of current range filtering methods. We have clearly pointed this out in \textcolor{blue}{Section 5.2.2} of the "Discussion" part of the paper and listed it as an important challenge for future work. Thank you for highlighting this point, which further emphasizes the significance of our work.
To make the paper more readable, we have added a more detailed explanation of the range attribute in Section 4.3.2.

In response to the large difference in index construction time shown in Figure 12a, we have revised the paper based on your suggestion. We added more detailed and in-depth analysis to provide a clearer explanation.(Please refer to the "Index construction time" part of Section 4.3.1, highlighted in blue in the revised manuscript)

Finally, your observation about the missing peak memory usage data for Milvus in Figure 12c is very insightful. Initially, we hesitated to include the peak memory usage of databases (such as Milvus) because they are full systems that run continuously and contain many components. This makes them quite different from other algorithms.

However, we agree with your point. To provide a more comprehensive evaluation, we have added the peak memory usage data of the database systems  in the revised version of the paper. This offers a more complete comparative view.\textbf{(Please refer to the "Build peak memory" part of Section 4.2.1 and  4.3.1, highlighted in blue in the revised manuscript)}

### 2025-07-28 VLDB2026/main.tex
change

\textit{\textbf{Index construction time.}}
\textcolor{orange}{Among the six regular-scale datasets shown in the left part of Figure~\ref{fig:build_time_comparison_query1}, the ones with longer build times are mainly PASE, VBASE, and ACORN-$\gamma$. PASE and VBASE only support single-threaded builds, and PASE is the slowest to build because its on-demand loading and caching strategies frequently trigger disk I/O and page scheduling, which seriously affects efficiency; VBASE uses full-memory builds and is relatively faster. ACORN-$\gamma$  has a large computational overhead because it needs to expand the neighbor list and evaluate a large number of candidates. The build time on some datasets even exceeds that of single-threaded VBASE. }

@@

	\textit{Single-Thread Search.}  
	\textcolor{violet}{As shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}, under near-perfect recall (recall approaching 1) conditions,} search often requires traversing larger candidate sets or deeper graph paths. This leads to increased computational cost and a sharp QPS drop for most methods. Despite this, UNG maintains high QPS due to its LNG structure, which eliminates unnecessary online filtering of irrelevant attributes, thereby improving query efficiency. \textcolor{violet}{Puck and CAPS perform relatively well on some datasets. Partition-based methods have limitations regarding different vector data distributions, and how to configure different partitions for different datasets remains a challenge. However, this also demonstrates that IVF-based methods are still very promising for hybrid queries.}

@@

	{\color{blue}{
		Algorithms that are highly sensitive to attribute distribution tend to rely passively on the statistical characteristics of the attributes in their design. For instance, graph-based indexing methods like FilteredVamana and NHQ are prone to degraded connectivity or distorted structures under power-law distributions, where label frequencies are highly imbalanced. Similarly, VBASE, which employs a post-filtering strategy, shows efficiency that directly correlates with attribute selectivity—the more selective the filter, the more computational resources are wasted on retrieving irrelevant vectors. Faiss, which searches within the nearest clusters, may also suffer when relevant data points are unevenly distributed across clusters due to attribute skew, potentially excluding qualified results.

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.}
	% Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, the range filtering experiment only uses one attribute, and the dataset is sorted in ascending order according to the attribute value size. WIT-Real and YT-Audio-Real use real attribute values, and the rest use randomly generated attribute values.}
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.} 
	Based on this definition, we evaluate algorithm performance under four query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. Due to the limitations of existing range filtering algorithms, our experiment uses only a single attribute, with the dataset sorted in ascending order by value. The WIT-Real and YT-Audio-Real datasets use real attribute values, while the rest use randomly generated ones. For space considerations, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. The experimental results show consistent performance across all datasets and query ranges.

@@

			\textbf{Large-Scale Dataset} & H & H & L & L &\textbf{H} &H  &M  &H  &M  &H  &\textbf{H}  &M/L  &L/M  &L/L  &M/M  &-  \\
			\textbf{Rare Attribute \& Low Range} &M  &M  &H  &H &\textbf{H}  &H  &H  &H  &L  &M  &\textbf{H} &M/M  &M/M  &L/L  &L/L  &L/L  \\
			\textbf{Robust to MultiModal Dataset} &H  &H  &L  &H  &\textbf{H}  &L  &M  &H  &M  &H  &\textbf{H}  &M/M  &M/M  &L/L  &L/L  &-/-  \\
			\textbf{Robust to Attribute Distribution} &L  &L  &H  &H  &H  &H  &-  &-  &-  &-  &-  &H/-  &L/-  &\textbf{H}/-  &L/-  &L/-  \\
			\textbf{High LID Dataset} &H  &M  &H  &M  &\textbf{H}  &M  &M  &H  &M  &H  &\textbf{H}  &L/L  &M/H  &L/L  &L/M  &L/L  \\
			\textbf{General-Purpose} &\faCheck  &\faCheck  &\faTimes  &\faTimes  &\faCheck  &\faTimes  &\faTimes  &\faTimes &\faTimes  & \faCheck &\faCheck&\faTimes  &\faCheck  &\faCheck  &\faCheck&\faTimes  \\
			\hline
		\end{tabular}
	}
	\centering
\end{table*}

@@

	\textcolor{orange}{
	iRange has relatively low time and space costs for index construction; however, its search performance declines significantly with increasing query range. Despite offering threading parameters for indexing, it does not effectively utilize multi-core parallelism in practical scenarios.
	SeRF incurs low time and space overhead but achieves only moderate overall query performance, with notably weaker results in small-range query scenarios.
	DSG supports unordered dynamic insertions and range queries, but its index construction is limited to single-threaded execution, resulting in long build times and poor practical applicability.}

		\textcolor{orange}{
	UNIFY adopts multiple filtering strategies tailored to different query range sizes, avoiding the limitations of a single strategy in specific scenarios. It requires only one index to support all queries, allows dynamic insertions, and features low time and space overhead during index construction, making it a more general-purpose solution.}
%	Table~\ref{tab:algo_scenarios} also shows the application scenarios of range filtering. iRange performs well on large-scale datasets, combining efficient construction and excellent query capabilities (S1). Faiss performs well in fast index construction (S2). When processing hard datasets, WinFilter, iRange, and DSG are recommended (S3), while WinFilter and iRange also perform well on easy datasets (S4). For general applications, DSG supports dynamic updates, while UNIFY provides flexible filtering strategies (S5). Faiss and SeRF show superior memory efficiency (S6). For static queries with high recall, WinFilter is the best choice (S7). In addition, WinFilter also performs well in OOD scenarios (S8).

	\subsubsection{\textbf{General Algorithm}}
\textcolor{orange}{All of the following methods support both attribute filtering and range filtering.}

\textcolor{orange}{
ACORN demonstrates average overall performance. Among its variants, ACORN-1 shows significant advantages in indexing speed on billion-scale datasets, but suffers from high memory consumption and delivers only moderate performance.}

>>>>

\textit{\textbf{Index construction time.}}
\textcolor{orange}{Among the six regular-scale datasets shown in the left part of Figure~\ref{fig:build_time_comparison_query1}, the ones with longer build times are mainly PASE, VBASE, and ACORN. PASE and VBASE only support single-threaded builds, and PASE is the slowest to build because its on-demand loading and caching strategies frequently trigger disk I/O and page scheduling, which seriously affects efficiency; VBASE uses full-memory builds and is relatively faster. ACORN-$\gamma$  has a large computational overhead because it needs to expand the neighbor list and evaluate a large number of candidates. The build time on some datasets even exceeds that of single-threaded VBASE. }

@@

	\textit{Single-Thread Search.}  
	\textcolor{violet}{As shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}, under near-perfect recall (recall approaching 1) conditions,} search often requires traversing larger candidate sets or deeper graph paths. This leads to increased computational cost and a sharp QPS (queries per second) drop for most methods. Despite this, UNG maintains high QPS due to its LNG structure, which eliminates unnecessary online filtering of irrelevant attributes, thereby improving query efficiency. \textcolor{violet}{Puck and CAPS perform relatively well on some datasets. Partition-based methods have limitations regarding different vector data distributions, and how to configure different partitions for different datasets remains a challenge. However, this also demonstrates that IVF-based methods are still very promising for hybrid queries.}

@@

	{\color{blue}{
		Algorithms that are highly sensitive to attribute distribution tend to rely passively on the statistical characteristics of the attributes in their design. For instance, graph-based indexing methods like FilteredVamana and NHQ are prone to degraded connectivity or distorted structures under power-law distributions, where label frequencies are highly imbalanced. Similarly, VBASE, which employs a post-filtering strategy, shows efficiency that directly correlates with attribute selectivity—the more selective the filter, the more computational resources are wasted on retrieving irrelevant vectors. FAISS-IVF, which searches within the nearest clusters, may also suffer when relevant data points are unevenly distributed across clusters due to attribute skew, potentially excluding qualified results.

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.}
	% Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, the range filtering experiment only uses one attribute, and the dataset is sorted in ascending order according to the attribute value size. WIT-Real and YT-Audio-Real use real attribute values, and the rest use randomly generated attribute values.}
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.} 
	Based on this definition, we evaluate algorithm performance under four query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, our experiment uses only a single attribute, with the dataset sorted in ascending order by value. The WIT-Real and YT-Audio-Real datasets use real attribute values, while the rest use randomly generated ones. }For space considerations, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. The experimental results show consistent performance across all datasets and query ranges.

@@

			\textbf{Large-Scale Dataset} & H & H & L & L &\textbf{H} &H  &M  &H  &M  &H  &\textbf{H}  &M/L  &L/M  &L/L  &M/M  &-  \\
			\textbf{Rare Attribute \& Low Range} &M  &M  &H  &H &\textbf{H}  &H  &H  &H  &L  &M  &\textbf{H} &M/M  &M/M  &L/L  &L/L  &L/L  \\
			\textbf{Robust to MultiModal Dataset} &H  &H  &L  &H  &\textbf{H}  &L  &M  &H  &M  &H  &\textbf{H}  &H/H  &M/M  &L/L  &L/L  &-/-  \\
			\textbf{Robust to Attribute Distribution} &L  &L  &H  &H  &H  &H  &-  &-  &-  &-  &-  &H/-  &L/-  &\textbf{H}/-  &L/-  &L/-  \\
			\textbf{High LID Dataset} &H  &M  &H  &M  &\textbf{H}  &M  &M  &H  &M  &H  &\textbf{H}  &L/L  &M/H  &L/L  &L/M  &L/L  \\
			\textbf{General-Purpose} &\faCheck  &\faCheck  &\faTimes  &\faTimes  &\faCheck  &\faTimes  &\faTimes  &\faTimes &\faTimes  & \faCheck &\faCheck&\faTimes  &\faCheck  &\faCheck  &\faCheck&\faTimes  \\
			\hline
		\end{tabular}
	}
	\centering
\end{table*}

@@

	\textcolor{orange}{
	iRange has relatively low time and space costs for index construction; however, its search performance declines significantly with increasing query range. Despite offering threading parameters for indexing, it does not effectively utilize multi-core parallelism in practical scenarios.
	SeRF incurs low time and space overhead but achieves only moderate overall query performance, with notably weaker results in small-range query scenarios.
	DSG supports unordered dynamic insertions and range queries, but its index construction is limited to single-threaded execution, resulting in long build times and poor practical applicability.
	UNIFY adopts multiple filtering strategies tailored to different query range sizes, avoiding the limitations of a single strategy in specific scenarios. It requires only one index to support all queries, allows dynamic insertions, and features low time and space overhead during index construction, making it a more general-purpose solution.}
%	Table~\ref{tab:algo_scenarios} also shows the application scenarios of range filtering. iRange performs well on large-scale datasets, combining efficient construction and excellent query capabilities (S1). Faiss performs well in fast index construction (S2). When processing hard datasets, WinFilter, iRange, and DSG are recommended (S3), while WinFilter and iRange also perform well on easy datasets (S4). For general applications, DSG supports dynamic updates, while UNIFY provides flexible filtering strategies (S5). Faiss and SeRF show superior memory efficiency (S6). For static queries with high recall, WinFilter is the best choice (S7). In addition, WinFilter also performs well in OOD scenarios (S8).

	\subsubsection{\textbf{General Algorithm}}
\textcolor{orange}{ACORN requires a separate index construction. Among its variants, ACORN-1 shows significant advantages in indexing speed on billion-scale datasets, but suffers from high memory consumption and delivers only moderate performance.}
